---
Nomaan Khan (nak150230)
CS 4375-501
Project 1
---

## Data Set 1 (ds1) Despcription:
    
    My first data set is a collection of data gathered about a bike sharing program in Portugal. It has data abou the number of
    casual and registered users each day for two years, it also has data about the weather, temperature, humidity, windspeed,
    and whether that day was a holiday. There are 16 columns and all are described below.
    
    My target variable is cnt which is total number of bikes used which includes both registered and casual users.
    I want to predict the number of users depending on various attributes, this prediction can be useful for the company
    to calculate the number of bikes they need on any particular day.
    
    I have performed linear regression and kNN classification on the data set.
    
    Number of Instances: 731
    Number of Attributes: 16 (including target)
    
    Attribute Description:
    
      1. - instant: record index
    	2. - dteday : date
    	3. - season : season (1:spring, 2:summer, 3:fall, 4:winter)
    	4. - yr : year (0: 2011, 1:2012)
    	5. - mnth : month ( 1 to 12)
    	6. - holiday : weather day is holiday or not (extracted from http://dchr.dc.gov/page/holiday-schedule)
    	7. - weekday : day of the week
    	8. - workingday : if day is neither weekend nor holiday is 1, otherwise is 0.
    	9. + weathersit : 
    		 - 1: Clear, Few clouds, Partly cloudy, Partly cloudy
    		 - 2: Mist + Cloudy, Mist + Broken clouds, Mist + Few clouds, Mist
    		 - 3: Light Snow, Light Rain + Thunderstorm + Scattered clouds, Light Rain + Scattered clouds
    		 - 4: Heavy Rain + Ice Pallets + Thunderstorm + Mist, Snow + Fog
    	10. - temp : Normalized temperature in Celsius. The values are divided to 41 (max)
    	11. - atemp: Normalized feeling temperature in Celsius. The values are divided to 50 (max)
    	12. - hum: Normalized humidity. The values are divided to 100 (max)
    	13. - windspeed: Normalized wind speed. The values are divided to 67 (max)
    	14. - casual: count of casual users
    	15. - registered: count of registered users
    	16. - cnt: count of total rental bikes including both casual and registered
    	
    	The data set is obtained from "https://archive.ics.uci.edu/ml/machine-learning-databases/00275/"
    	You have to download the zip file and use the day.csv file.
      For full citation see the citation section at the end.

```{r}
ds1 <- read.table("C:/Users/ADMIN/Desktop/Academia 2.0/Junior/Spring 2018/Machine Learning/HW/Project 1/day.csv", header=TRUE, sep=",")
```


Using R functions on the data set.

```{r}
names(ds1)
head(ds1)
str(ds1)
dim(ds1)
summary(ds1)
```

```{r}
par(mfrow=c(1,2))
plot(as.factor(ds1$weathersit),ds1$cnt, ylab = "Total Users", xlab = "Weather Situation")
plot(ds1$atemp,ds1$cnt, ylab = "Total Users", xlab = "Normalized Temperature")
```
From the first plot we can see that more users on a clear day.
From the second plot we see that as the normalized temperature rises the count of users tends to rise. 

Converting to factors, because all these variable have integer values which represent characters.

```{r}
ds1$season <- as.factor(ds1$season)
ds1$mnth <- as.factor(ds1$mnth)
ds1$weekday <- as.factor(ds1$weekday)
ds1$workingday <- as.factor(ds1$workingday)
ds1$weathersit <- as.factor(ds1$weathersit)
```

Splitting the data randomly into 75% train and 25% test.

```{r}
set.seed(1234)
i <- sample(1:nrow(ds1), nrow(ds1)*0.75, replace=FALSE)
train <- ds1[i,]
test <- ds1[-i,]
```


Building a linear model, where the target is total number of casual users and the predictors are humidity, average
temperature, weather situation, season and wind speed.

```{r}
lm1 <- lm(cnt~atemp*hum*temp+atemp+temp+hum+mnth+weathersit+season, data=train)
pred <- predict(lm1, newdata=test)
cor(pred, test$cnt)
```

```{r}
summary(lm1)
```


Calculating MSE and RSE.

```{r}
mse <- mean(lm1$residuals^2)
mse
rse <- sqrt(mse)
rse
```

Concverting the following attributes to numeric so that they can be used for kNN regression.

```{r}
ds1$weathersit <- as.integer(ds1$weathersit)
ds1$mnth <- as.integer(ds1$mnth)
ds1$season <- as.integer(ds1$season)
```

Perfoming kNN regression on scaled data. The predictors are the same as the ones in the linear model.

```{r}
library('caret')
df <- data.frame(scale(ds1[,c(3,9,10,11,12,13,16)])) # scaling total number of casual users, humidity, average temperature,                                                         # weather situation, season, mnth and wind speed.
train2 <- df[i,]
test2 <- df[-i,]

fit <- knnreg(train2[,1:6],train2[,7],k=12)
predictions <- predict(fit, test2[,1:6])
cor(predictions, test2$cnt)                          # calculating accuracy
```

```{r}
summary(fit)

r_squared <- 1 - sum((test2$cnt-predictions)^2)/sum((test2$cnt-mean(test2$cnt))^2)
r_squared
```

Summary doesn't tell us much, but R squared will.

## Analysing Models/Algorithm Results:
    
    Both linear model and kNN regression have good accuracy and good R squared.
    
    The linear model has an accuracy of 0.7462853 and R squared of 0.613. This is high accuracy and the model is able to
    explain 61.3% of the variabiity. It also has a low p value which tells us that the predictors are predicting well.
    
    kNN regression gives an accuracy of 0.7902023 and R squared of 0.621735. This is high accuracy and the model is able to
    explain 62.1735% of the variabiity.
    
    kNN regression just edges the linear model in both accuracy and explaining variability, and is the best for this data set.
    
    


## Citations:
    
    [1] Contains link for data sets.
  
    1. Data Set 1 (Bike Sharing in Portugal)
    
        [1] Lichman, M. (2013). UCI Machine Learning Repository                                                                            [https://archive.ics.uci.edu/ml/machine-learning-databases/00275/]. Irvine, CA: University of California, School of             Information and Computer Science.
    
        [2] Fanaee-T, Hadi, and Gama, Joao, "Event labeling combining ensemble detectors and background knowledge", Progress in             Artificial Intelligence (2013): pp. 1-15, Springer Berlin Heidelberg, doi:10.1007/s13748-013-0040-3.
        